---
title: 'Global Terrorism: from 1970 to 2019'
author: "Riccardo Bertoli"
output:
  ioslides_presentation:
    css: main.css
    incremental: yes
editor_options: 
  chunk_output_type: inline    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, message = FALSE, warning = FALSE)
```
## <b>Where does the data come from?</b> 
<p>From a free open source database maintained by researchers at the National Consortium for the Study of Terrorism and Responses to Terrorism (START), headquartered at the University of Maryland. The data shows information that spans from 1970 until 2019.The last update was on the 17th September of 2019 and the are currently three versions of the data set. During the merging of the data to the servers the Consortium lost all 1993's data and other dates too. However they managed to recover all 1993's missing data so I combined the two sets.</p>
```{r echo=FALSE}
#librarys
library(readxl)
library(readr)
library(tibble)
library(magrittr)
library(dplyr)
library(ggplot2) #graphics
library(plotly) #graphics
library(fmsb) #for radar graphics
library(wordcloud)#for wordcloud graphics
library(treemap)#fro treempas graphics
#read the data
terrorism <-read_excel("gtd1970_2019update.xlsx",1)
#system.time(terrorism <-read_excel("gtd1970_2019update.xlsx",1))
dataset <- as_tibble(terrorism)
#View(dataset)
```
```{r echo=FALSE}
#what I'm focusing on
#dataset with kills
maink <-(dataset %>%
        filter(nkill>0)%>%   
        select(iyear,imonth,iday,country_txt,region_txt,latitude,longitude,provstate,city,attacktype1_txt,targtype1_txt,target1,gname,weaptype1_txt,nkill,nkillus))
#dataset without kills
main<-(dataset %>%
      filter(nkill==0)%>%
      select(iyear,imonth,iday,country_txt,region_txt,latitude,longitude,provstate,city,attacktype1_txt,targtype1_txt,target1,gname,weaptype1_txt,nkill))
#maink
#main
```
## Global Terrorism
<p><b>The World situation seen by different points of view:</b></p>
<p>A Globe map</p>
```{r echo=FALSE}
#world situation
world <- map_data("world")
#global terrorism graphic
ggplot() + geom_polygon(data = world, aes(x=long, y = lat, group = group)) + coord_fixed(1.3)+geom_point(data=maink,aes(x=longitude, y=latitude),color="red",size=0.5)+geom_point(data=main,aes(x=longitude, y=latitude),color="yellow",size=0.001)+labs(title="Global Terrorism")
```
<p>Two different graphics (function and barplot) showing the casualties through the years</p>
```{r echo=FALSE}
#worlddeath barplot/frequency
#per group iyear
years <- ( maink %>%
           group_by(iyear) %>%
           summarise(nkill=sum(nkill,na.rm = TRUE)))
fig1<-plot_ly(years, x = ~iyear,y=~nkill,name="Victims During The Years" ,type = 'scatter', mode = 'lines')
fig1 <-fig1 %>% layout(title="Victims During The Years",xaxis=list(title="Regions"),yaxis=list(title="Victims"))
fig1
ggplot(maink, aes(fill=region_txt, y=nkill, x=iyear)) + geom_bar(position="stack", stat="identity",width = 0.8)+labs(title="Victims Per Regions During The Years",x="Years",y="Victims",fill="Regions")
```
<p>Focusing on the regions</p>
```{r echo=FALSE}
#per group
gt <- (maink %>%
          group_by(region_txt) %>%
          summarise(nkill=sum(nkill,na.rm=TRUE)))
#world victims barplot
fig2<-plot_ly(gt,x=~region_txt,y=~nkill,name="Victims per Region",color=I("red"),type="bar")
fig2<-fig2 %>% layout(title="Victims Per Region",xaxis=list(title="Regions"),yaxis=list(title="Victims"))
fig2
```

## Middle East & North Africa
<b>A simple wordcloud to show which state is suffering the most.</b>
```{r echo=FALSE}
#single region dataset
mena<-( dataset %>%
             filter(region_txt=="Middle East & North Africa") %>%   
             select(iyear,imonth,iday,country_txt,region_txt,latitude,longitude,provstate,city,attacktype1_txt,targtype1_txt,target1,gname,weaptype1_txt,nkill,nkillus)) 
#mena
#per group
menav <- (mena %>%
          group_by(country_txt) %>%
          summarise(nkill=sum(nkill,na.rm=TRUE),nkillus=sum(nkillus,na.rm=TRUE)))
#menav
wordcloud(words = menav$country_txt, freq = menav$nkill, min.freq = 0, random.order=FALSE, rot.per=0.35,colors=brewer.pal(8, "Dark2"))
```

## IRAQ'S WARS
<p>Iraq suffered more than 80.000 victims through the years. In the last twenty years they suffered two wars (during the 2003-2011 invasion and the ISIS proclamation of 2013-2017)</p>
<br>
```{r echo=FALSE}
#four most suffering states
mena_principal<-(mena %>%
                 filter(country_txt=="Iraq"| country_txt=="Algeria" | country_txt=="Syria" | country_txt=="Yemen") %>%   
                 select(iyear,imonth,iday,country_txt,region_txt,latitude,longitude,provstate,city,attacktype1_txt,targtype1_txt,target1,gname,weaptype1_txt,nkill,nkillus))
#mena_principal
#iraq dataset focusing on the victims
iraq<-(maink %>%
        filter(country_txt=="Iraq") %>%   
        select(iyear,imonth,iday,country_txt,region_txt,latitude,longitude,provstate,city,attacktype1_txt,targtype1_txt,target1,gname,weaptype1_txt,nkill,nkillus))
#the three wars
iraqusawar<-(iraq %>%
           group_by(iyear) %>%
           summarise(nkill = sum(nkill,na.rm=TRUE),nkillus=sum(nkillus,na.rm = TRUE)) %>%
           filter(iyear>=2003 & iyear<=2011))
isisus<-(iraq %>%
           group_by(iyear) %>%
           summarise(nkill = sum(nkill,na.rm=TRUE),nkillus=sum(nkillus,na.rm = TRUE)) %>%
           filter(iyear>=2011 & iyear<=2017))
#add column war in the two datasets
iraqusawar <-cbind(iraqusawar,war="Iraq War")
isisus <-cbind(isisus,war="ISIS War")
#sum of usa soldiers
#sum(iraqusawar$nkillus,na.rm = TRUE)=255
#sum(isisus$nkillus,na.rm = TRUE)=12
#iraqusawar
#isisus
#combine two datasets
iraqwar <-rbind(iraqusawar,isisus)
#victims per war
#aggregate(nkill~war,iraqusawar,sum)
#aggregate(nkill~war,isisus,sum)
#iraqwar
# Compute the position of labels
iraqwar <- iraqwar %>% 
  arrange(desc(war)) %>%
  mutate(prop = nkill / sum(iraqwar$nkill) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )
#piechart
ggplot(iraqwar, aes(x="", y=prop, fill=war)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() + 
  theme(legend.position="bottom") +
  geom_text(aes(y = ypos, label = nkill), color = "black", size=2.5,nudge_x = 0.6,nudge_y = 0.5) +
  scale_fill_brewer(palette="Set1") +
  labs(title="Comparing the two wars",fill="Type of War:")
```
<ul>These two recent conflicts brought:
  <li>26.616 civilian victims in the IRAQ war plus 255 USA soldiers.</li>
  <li>53.336 civilian victims in the ISIS war plus 12 USA soldiers.</li>
  <li>During these wars attacks 79.952 lives were lost.</li>
</ul>
## How did they die? Who killed these people? Where did these attacks happen?
<p>The main weapons used in the attacks</p>
```{r echo=FALSE}
#filter per year and group per weapon and year
iraqweapons <- (iraq %>%
          filter(iyear>=2003 & nkill>=10)%>%
          group_by(iyear,weaptype1_txt) %>%
          summarise(nkill=sum(nkill,na.rm = TRUE)))
#weapons graphic
figw <- plot_ly(iraqweapons, x = ~iyear, y = ~nkill, color=~weaptype1_txt,colors = "Set1",type = 'bar')
figw<-figw %>% layout(title="Victims Per Weapon",xaxis=list(title="Years"),yaxis=list(title="Victims"))
figw
```
<p>Here are the main terrorist groups</p> 
```{r echo=FALSE}
#iraq
#per groups
iraqgroups<-(iraq %>%
              filter(nkill>=50 & iyear>=2003) %>%
              group_by(iyear,gname) %>%
              summarise(nkill=sum(nkill,na.rm=TRUE)) %>%
              arrange(-iyear))
#iraqgroups
#sum(iraqgroups$nkill,na.rm=TRUE) = 79760
fig3 <- plot_ly(iraqgroups, x = ~iyear,y=~nkill,color = ~gname,colors = "Set1")
fig3<-fig3 %>% layout(title="Victims Per Groups",xaxis=list(title="Years"),yaxis=list(title="Victims"))
fig3
```
<p>Cities that had more than 100 victims.</p>
```{r echo=FALSE}
#locations
#per groups iyear,latitude,longitude,city,nkill
iraqcity<-(iraq%>%
              group_by(iyear,city,latitude,longitude)%>%
              summarise(nkill=sum(nkill,na.rm=TRUE))%>%
              filter(nkill>=100)%>%
              select(iyear,latitude,longitude,city,nkill)%>%
              arrange(-iyear))
#iraqcity
#iraq map cities
figiraq <- iraqcity 
figiraq <- figiraq %>%
  plot_ly(
    lat = ~latitude,
    lon = ~longitude,
    marker = list(color = "red"),
    type = 'scattermapbox',
    hovertext=~city) 
figiraq <- figiraq %>%
  layout(
    mapbox = list(
      style = 'open-street-map',
      zoom =5,
      center = list(lon = 43.679291, lat = 33.223191)))
figiraq
```
<ul>
  <p>These are the primary cities and towns have been hit by terrorism:</p>
  <li>Baghdad has reached 21.394 fatalities.</li>
  <li>Mosul has reached 7.279 fatalities.</li>
</ul>
```{r echo=FALSE}
#how terroris grow in the citys
fig4 <- plot_ly(iraqcity, x = ~iyear,y=~nkill,color = ~city,colors = "Set1")
fig4<-fig4 %>% layout(title="Victims Per City During the Years",xaxis=list(title="Years"),yaxis=list(title="Victims"))
fig4
#sum of single city
iraqcity<-(iraq%>%
              group_by(city)%>%
              summarise(nkill=sum(nkill,na.rm=TRUE))%>%
              filter(nkill>=600))
iraqcityfig <- plot_ly(iraqcity,type = 'scatterpolar',r = ~nkill,theta = ~city,fill = 'toself',name="Victims Per City/Town") 
iraqcityfig<- iraqcityfig %>% layout(polar = list(radialaxis = list(visible = T,range = c(0,5000))),showlegend = FALSE,(title="Victims Per City/Town"))
iraqcityfig
```

## South Asia
```{r echo=FALSE}
#single region dataset
sa<-( dataset %>%
             filter(region_txt=="South Asia") %>%   
             select(iyear,imonth,iday,country_txt,region_txt,latitude,longitude,provstate,city,attacktype1_txt,targtype1_txt,target1,gname,weaptype1_txt,nkill,nkillus)) 
#sa
#per gname
sav <- (sa %>%
          group_by(country_txt) %>%
          summarise(nkill=sum(nkill,na.rm=TRUE),nkillus=sum(nkillus,na.rm=TRUE)))
#sav
#menav
wordcloud(words = sav$country_txt, freq = sav$nkill, min.freq = 0, random.order=FALSE, rot.per=0.35,colors=brewer.pal(8, "Dark2"))
```

## Afghanistan, India, Sri Lanka and Pakistan
<p>These four states, as we saw before, are the most suffering.<p>
```{r echo=FALSE}
sav <-(sav %>% filter(nkill>10000))
fig5<-plot_ly(sav,x=~country_txt,y=~nkill,name="Victims per Country",color=I("red"),type="bar")
fig5<-fig5 %>% layout(title="Victims per Country",xaxis=list(title="Regions"),yaxis=list(title="Victims"))
fig5
```
## How did they die? Who killed these people? Where did these attacks happen?
<p>The main weapons used in the attacks.</p>
```{r echo=FALSE}
#weapons
saweapons <- (sa %>%
                filter(nkill>=10)%>%
                group_by(iyear,weaptype1_txt) %>%
                summarise(nkill=sum(nkill,na.rm = TRUE)))
#saweapons
#weapons
saw <- plot_ly(saweapons, x = ~iyear, y = ~nkill, color=~weaptype1_txt,colors = "Set1",type = 'bar')
saw<-saw %>% layout(title="Victims Per Weapon",xaxis=list(title="Years"),yaxis=list(title="Victims"))
saw
```
<p>Here are the main terrorist groups.</p>
```{r echo=FALSE}
#per gname
sagroups<-(sa %>%
              group_by(iyear,gname) %>%
              summarise(nkill=sum(nkill,na.rm=TRUE)) %>%
              filter(nkill>=100) %>%
              arrange(-nkill))
#sagroups
fig5 <- plot_ly(data = sagroups, x = ~iyear, y = ~nkill, color = ~gname,colors="Set1")
fig5 <- fig5 %>% layout(title="Victims Per Group",xaxis=list(title="Years"),yaxis=list(title="Victims"))
fig5
```
<p>Cityies that had more than 100 victims.</p>
```{r echo=FALSE}
#location
sacity <-(sa %>%
          group_by(iyear,city,latitude,longitude)%>%
              summarise(nkill=sum(nkill,na.rm=TRUE))%>%
              filter(nkill>=100)%>%
              select(iyear,latitude,longitude,city,nkill)%>%
              arrange(-iyear))
#sacity
#south asia map
figsa <- sacity 
figsa <- figsa %>%
  plot_ly(
    lat = ~latitude,
    lon = ~longitude,
    marker = list(color = "red"),
    type = 'scattermapbox',
    hovertext = sacity$city)
figsa <- figsa %>%
  layout(
    mapbox = list(
      style = 'open-street-map',
      zoom =3,
      center = list(lon = 67.709953, lat = 33.93911)))
figsa
```
<p>These are the primary cities and towns hit by terrorism.</p>
```{r echo=FALSE}
#how terroris grow in the citys
fig6 <- plot_ly(sacity, x = ~iyear,y=~nkill,color = ~city,colors = "Set1")
fig6<-fig6 %>% layout(title="Victims Per City During the Years",xaxis=list(title="Years"),yaxis=list(title="Victims"))
fig6
#sum of single city
sacityg<-(sa%>%
              group_by(city)%>%
              summarise(nkill=sum(nkill,na.rm=TRUE))%>%
              filter(nkill>=600))
fig7 <- plot_ly(sacityg,type = 'scatterpolar',r = ~nkill,theta = ~city,fill = 'toself',name="Victims Per City/Town") 
fig7<- fig7 %>% layout(polar = list(radialaxis = list(visible = T,range = c(0,5000))),showlegend = FALSE,(title="Victims Per City/Town"))
fig7
```
## Sub-Saharan Africa
```{r echo=FALSE}
#single region dataset
ssa<-(dataset %>%
             filter(region_txt=="Sub-Saharan Africa") %>%   
             select(iyear,imonth,iday,country_txt,region_txt,latitude,longitude,provstate,city,attacktype1_txt,targtype1_txt,target1,gname,weaptype1_txt,nkill,nkillus))
#ssa
#per country
ssav <- (ssa %>%
          group_by(country_txt) %>%
          summarise(nkill=sum(nkill,na.rm=TRUE),nkillus=sum(nkillus,na.rm=TRUE))%>%
          arrange(country_txt))
#ssav
#menav
wordcloud(words =ssav$country_txt, freq = ssav$nkill, min.freq = 0, random.order=FALSE, rot.per=0.35,colors=brewer.pal(8, "Dark2"))
```

## Burundi, Democratic Republic of the Congo, Nigeria, Somalia and Sudan 
```{r echo=FALSE}
ssav <-(ssav %>% filter(nkill>=3980))
fig8<-plot_ly(ssav,x=~country_txt,y=~nkill,name="Victims per Country",color=I("red"),type="bar")
fig8<-fig8 %>% layout(title="Victims per Country",xaxis=list(title="Regions"),yaxis=list(title="Victims"))
fig8
```
## How did they die? Who killed these people? Where did these attacks happen?
<p>The main weapons used in the attacks.</p>
```{r echo=FALSE}
#weapons
ssaweapons <- (ssa %>%
                filter(nkill>0)%>%
                group_by(iyear,weaptype1_txt) %>%
                summarise(nkill=sum(nkill,na.rm = TRUE)))
#ssaweapons
#weapons
ssaw <- plot_ly(ssaweapons, x = ~iyear, y = ~nkill, color=~weaptype1_txt,colors = "Set1",type = 'bar')
ssaw<-ssaw %>% layout(title="Victims Per Weapon",xaxis=list(title="Years"),yaxis=list(title="Victims"))
ssaw
```
<p>Here are the principal terrorist groups.</p>
```{r echo=FALSE}
#per gname
ssagroups<-(ssa %>%
              group_by(iyear,gname) %>%
              summarise(nkill=sum(nkill,na.rm=TRUE)) %>%
              filter(nkill>=200) %>%
              arrange(-nkill))
#ssagroups
fig8 <- plot_ly(data = ssagroups, x = ~iyear, y = ~nkill, color = ~gname, colors = "Set1")
fig8 <- fig8 %>% layout(title="Victims Per Group",xaxis=list(title="Years"),yaxis=list(title="Victims"))
fig8
```
<p>Cities and towns with more than 50 victims.</p>
```{r echo=FALSE}
#location
ssacity <-(ssa %>%
          group_by(iyear,city,latitude,longitude)%>%
              summarise(nkill=sum(nkill,na.rm=TRUE))%>%
              filter(nkill>=50)%>%
              select(iyear,latitude,longitude,city,nkill)%>%
              arrange(-iyear))
#ssacity
#sub-sharan africa map
figssa <- ssacity
figssa <- figssa %>%
  plot_ly(
    lat = ~latitude,
    lon = ~longitude,
    marker = list(color = "red"),
    type = 'scattermapbox',
    hovertext = ssacity$city) 
figssa <- figssa %>%
  layout(
    mapbox = list(
      style = 'open-street-map',
      zoom =2,
      center = list(lon = 8.675277, lat = 9.081999)))
figssa
```
<p>These are the primary cities and towns hit by terrorism.</p>
```{r echo=FALSE}
ssacity <- ssacity %>% filter(nkill>=200)
fig9 <- plot_ly(ssacity, x = ~iyear,y=~nkill,color = ~city,colors = "Set1")
fig9<-fig9 %>% layout(title="Victims Per City During the Years",xaxis=list(title="Years"),yaxis=list(title="Victims"))
fig9
#sum of single city 5932
ssacityg<-(ssa%>%
              group_by(city)%>%
              summarise(nkill=sum(nkill,na.rm=TRUE))%>%
              filter(nkill>=500))
fig10 <- plot_ly(ssacityg,type = 'scatterpolar',r = ~nkill,theta = ~city,fill = 'toself',name="Victims Per City/Town") 
fig10<- fig10 %>% layout(polar = list(radialaxis = list(visible = T,range = c(0,5000))),showlegend = FALSE,(title="Victims Per City/Town"))
fig10
```
## Which is the state that suffered the most in a single day? Which is the city that sufferd the most during the years?
<p>The deadliest day for a single terrorist attack.</p>
```{r echo=FALSE}
#worstday
worst_day <- (maink %>%
              filter(nkill>=500) %>%
              group_by(iday,imonth,iyear,city)%>%
              summarise(nkill=sum(nkill,na.rm = TRUE))%>%
              arrange(-nkill)
)
#worst_day
worst_fig <- plot_ly(worst_day,type = 'scatterpolar',r = ~nkill,theta = ~city,fill = 'toself') 
worst_fig<- worst_fig %>% layout(polar = list(radialaxis = list(visible = T,range = c(0,3000))),showlegend = F)
worst_fig
```
<p>The worst cities.</p>
```{r echo=FALSE}
#worstcity
worstcity <- (maink %>%
              group_by(city)%>%
              summarise(nkill=sum(nkill,na.rm = TRUE))%>%
              filter(nkill>=1000)%>%  
              arrange(-nkill))
#worstcity
city <- plot_ly(worstcity, x=~city,y=~nkill, type="bar", name="The Worst City",colors=I("red"))
city <- city %>% layout(title="The City With More Victims",xaxis=list(title="Years"),yaxis=list(title="Victims"))
city
```